---
title: "Can winter Rs a proxy for annual Rh? Can this be scaled globally?"
author: "Stephanie Pennington"
date: "9/29/2021"
output: html_document
---

**Hypothesis**: Annual soil respiration (Rs) is composed of autotropic (Ra) and heterotrophic (Rh) respiration. In the winter, trees and other rooted organisms are dormant and ideally produce a negligible amount of CO2 (responsible for Ra). Therefor we hypothesize that wintertime Rs is mainly comprised of microbial activity (Rh), and wintertime Rs can be used as a proxy for annual Rh. This relationship could allow for globally scaled estimates of Rh.

```{r setup, include = FALSE}
library(dplyr)
library(readr)
library(tidyr)
library(DT)

# We want to load in the SRDB and search for relevant studies
srdb_data <- read_csv("../Data/srdb-data.csv")

vars <- c("Rs_winter", "Rh_annual")
```

```{r , inclue = FALSE}
# How many Rs_winter observations are there in SRDB?
winter_rs <- nrow(drop_na(srdb_data, Rs_winter))
    
# Also how many Rs_winter observations do we have that also have an Rh_annual value? 
srdb_data %>% 
    select(Record_number, Study_number, Study_midyear, Biome, Rs_winter, Rs_annual, Rh_annual) %>% 
    drop_na(all_of(vars)) -> srdb_winter

rsrh_obs <- nrow(srdb_winter)
rsrh_studies <- length(unique(srdb_winter$Study_number))

# Let's make an interactive table of the winter SRDB entries
datatable(srdb_winter)
    
```

### There are `r winter_rs` `Rs_winter` measurements reported in the SRDB.

### There are `r rsrh_obs` observations in `r rsrh_studies` studies that report both `Rs_winter` and `Rh_annual` reported in the SRDB.

#### Questions
- What exactly is Rh composed of? Rr and what?
- How can I efficiently search for these papers?
