---
title: "Science Question: Can winter Rs a proxy for annual Rh? Can this be scaled globally?"
author: "Stephanie Pennington"
date: "9/29/2021"
output: html_document
---

Annual soil respiration (Rs) is composed of autotropic (Ra) and heterotrophic (Rh) respiration. 

**Hypothesis 1:** In the winter, plants are dormant and produce negligible respiration. Therefore we hypothesize that wintertime Rs is mainly comprised of microbial activity (Rh). 

_How do we test H1?_ 
- COSORE? Need sites with wintertime Rs and Rh measurements
- SRDB? What studies report Rs_winter and separate Rh and Ra annually. Some may have reported wintertime Rh but not entered into SRDB
- Lit searching?
- AmeriFlux? Sites report ER (total ecosystem respiration); if in winter ER gets close to Rs (from SRDB?) that's suggestive, needs more research
- transplant exp???

**Hypothesis 2:** Wintertime Rs (equals Rh, if H1 is true) can be used as a proxy for annual Rh. This relationship could allow for biome-scale estimates of Rh.

_How do we test H2?_
- SRDB: look at studies that report Rs_winter and Rh_annual
- COSORE?

```{r setup, include = FALSE}
library(dplyr)
library(readr)
library(tidyr)
library(DT)
library(ggplot2)
library(raster)
library(cowplot)
library(ggmap)
library(viridis)

theme_set(theme_minimal())

# We want to load in the SRDB and search for relevant studies
srdb_data <- read_csv("../Data/srdb-data.csv")

vars <- c("Rs_winter", "Rh_annual")
```

```{r data-prep, echo = FALSE}
# How many Rs_winter observations are there in SRDB?
winter_rs <- nrow(drop_na(srdb_data, Rs_winter))
    
# Also how many Rs_winter observations do we have that also have an Rh_annual value? 
srdb_data %>% 
  # Study 11897 had incorrect coordinates and was not matching in WorldClim data
  mutate(Latitude = if_else(Study_number == "11897", 27.46, Latitude),
         Longitude = if_else(Study_number == "11897", -81.75, Longitude)) %>% 
   # select(Record_number, Study_number, Study_midyear, Biome, Rs_winter, Rs_annual, Rh_annual) %>% 
    drop_na(all_of(vars)) -> srdb_winter

rsrh_obs <- nrow(srdb_winter)
rsrh_studies <- length(unique(srdb_winter$Study_number))

# Let's make an interactive table of the winter SRDB entries
srdb_winter %>% 
    dplyr::select(Record_number, Study_number, Study_midyear, Biome, Rs_winter, Rs_annual, Rh_annual) %>% 
    datatable()
    
```

### There are `r winter_rs` `Rs_winter` measurements reported in the SRDB.

### There are `r rsrh_obs` observations in `r rsrh_studies` studies that report both `Rs_winter` and `Rh_annual` reported in the SRDB.

#### Key Terms to Search For
- soil respiration; winter; heterotrophic respiration




Extract locations from srdb_winter
 Warmest 3 months - Coldest 3 months -> temperature seasonality
 Wettest 3 months - Dryest 3 months -> precipitation seasonality
 Compute MAT and MAP
 
```{r mat-map}
# Download worldclim data for precip and tmean if necessary, into wc10/ folder
precip <- getData("worldclim", path = "../Data/wc10/", var = "prec", res = 10, download = !file.exists("../Data/wc10/wc10/precip1.hdr"))
tmean <- getData("worldclim", path = "../Data/wc10/", var = "tmean", res = 10, download = !file.exists("../Data/wc10/wc10/tmean1.hdr"))

# Pull out srdb_winter dataset latitudes and longitudes
srdb_winter %>%
  dplyr::select(Study_number, Record_number, Longitude, Latitude) -> winter_coords

# Extract srdb_winter location data from worldclim data for precip...
raster::extract(precip, winter_coords[3:4]) -> precip_coords
apply(precip_coords, 1, sum) -> map_winter
cbind(winter_coords, map_winter) -> map_coords


# ...and tmean
raster::extract(tmean, winter_coords[3:4]) -> tmean_vals
apply(tmean_vals, 1, mean) -> mat_winter
cbind(map_coords, mat_winter)  %>%
  # Temp data is stored in degC * 10, so we need to divide to get back to degC
  mutate(mat_winter = mat_winter / 10)   -> winter_points
# %>% 
#     distinct(Latitude, Longitude, .keep_all = TRUE)

# Extract global climate space data
raster::as.data.frame(precip, xy = TRUE) %>%
  drop_na() -> precip_global

# Calculate annual sum for precip...
precip_global %>%
  dplyr::select(-x, -y) %>%
  apply(1, sum) -> map_global

raster::as.data.frame(tmean, xy = TRUE) %>%
  drop_na() -> tmean_global

# ...and mean for temperature
tmean_global %>%
  dplyr::select(-x, -y) %>%
  apply(1, mean) -> mat_global

# Create tibble with corresponding coordinates
tibble(x = tmean_global$x, y = tmean_global$y, mat = as.vector(mat_global)) -> mat
tibble(x = precip_global$x, y = precip_global$y, map = as.vector(map_global)) -> map

left_join(map, mat, by = c("x", "y")) %>%
  # Temp data is stored in degC * 10, so we need to divide to get back to degC
  mutate(mat = mat / 10) -> map_mat_global

# Try Mark Tjoelker's suggestion re Whittaker biomes
library(plotbiomes) # devtools::install_github("valentinitnelav/plotbiomes")
p_inset <- whittaker_base_plot() +
  geom_point(data = winter_points, aes(x = mat_winter, y = map_winter / 10),
             color = "black", shape = 4) +
  theme(axis.title = element_blank(),
        axis.text = element_text(size = 8),
        legend.text = element_text(size = 7),
        legend.key.size = unit(0.4, "lines"),
        legend.position = c(0.35, 0.75),
        legend.title = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white"),
        panel.border = element_rect(colour = "black", fill = NA, size = 0.5))

# SP's main climate space plot
p <- ggplot() +
  geom_hex(data = map_mat_global,
           aes(x = mat, y = map / 10), bins = 100, na.rm = TRUE) +
  scale_fill_viridis_c(name = "Grid cells", begin = 0.85, end = 0) +
  geom_point(data = winter_points, aes(x = mat_winter, y = map_winter / 10),
             color = "black", shape = 4, size = 1.5, na.rm = TRUE) +
  theme_minimal() +
  labs(x = "MAT (°C)", y = "MAP (cm)")
print(p)
#ggsave_quiet("figures/figure3-climate.png")

suppressMessages(library(cowplot, quietly = TRUE))
p_new <- ggdraw() +
  draw_plot(p) +
  draw_plot(p_inset, x = 0.1, y = 0.5, width = 0.4, height = 0.45)
print(p_new)
#save_plot("figures/figure3-climate-whittaker.png", p_new, base_height = 5)
```

```{r seasonality}

cbind(winter_coords, precip_coords) %>% 
pivot_longer(prec1:prec12) %>% 
  group_by(Record_number) -> precip_monthly

precip_monthly %>% 
  arrange(desc(value), .by_group = TRUE) %>% 
  slice(1:3) %>% 
  summarise(P_High_Avg = round(mean(value), digits = 2))  -> p_high

precip_monthly %>% 
  arrange(value, .by_group = TRUE) %>% 
  slice(1:3) %>% 
  summarise(P_Low_Avg = round(mean(value), digits = 2)) -> p_low 

cbind(winter_coords, tmean_vals) %>% 
pivot_longer(tmean1:tmean12) %>% 
  mutate(value = value / 10) %>% 
  group_by(Record_number) -> tmean_monthly

tmean_monthly %>% 
  arrange(value, .by_group = TRUE) %>% 
  slice(1:3) %>% 
  summarise(T_Low_Avg = round(mean(value), digits = 2)) -> t_low 

tmean_monthly %>% 
  arrange(desc(value), .by_group = TRUE) %>% 
  slice(1:3) %>% 
  summarise(T_High_Avg = round(mean(value), digits = 2)) -> t_high

p_high %>% 
  left_join(p_low, by = c("Record_number")) %>% 
  left_join(t_high, by = c("Record_number")) %>% 
  left_join(t_low, by = c("Record_number")) %>% 
  mutate(precip_seasonality = P_High_Avg - P_Low_Avg, temp_seasonality = T_High_Avg - T_Low_Avg) %>% 
  dplyr::select(Record_number, precip_seasonality, temp_seasonality) -> seasonality

srdb_winter %>% 
  left_join(seasonality, by = "Record_number") %>% 
    ggplot(aes(x = Rs_winter, y = Rh_annual)) +
    geom_point(aes(color = temp_seasonality))


```

Compute seasonality

```{r maps}

world <- map_data("world")
smap <- get_stamenmap(bbox = c(left = -180, bottom = -80, right = 179.9999, top = 85), zoom = 3)

seasonality %>% 
  left_join(winter_coords, by = "Record_number") -> t

# temp seasonality
  ggmap(smap) +
  geom_point(data = t, aes(x = Longitude, y = Latitude,
                           color = temp_seasonality)) +
  scale_color_viridis(begin = 1, end = 0.3, option = "plasma")
  
# precip seasonality  
  ggmap(smap) +
  geom_point(data = t, aes(x = Longitude, y = Latitude,
                           color = precip_seasonality)) +
  scale_color_viridis(begin = 1, end = 0.3, option = "viridis")

# map
  ggmap(smap) +
  geom_point(data = winter_points, aes(x = Longitude, y = Latitude,
                           color = map_winter)) +
  scale_color_viridis(begin = 1, end = 0.3, option = "mako")
  
#mat
  ggmap(smap) +
  geom_point(data = winter_points, aes(x = Longitude, y = Latitude,
                           color = mat_winter)) +
  scale_color_viridis(begin = 1, end = 0.3, option = "mako")
```

```{r rs-rh, echo = FALSE}

srdb_winter %>% 
  dplyr::select(Record_number, Study_number, Rs_winter, Rh_annual) %>% 
  left_join(seasonality, by = "Record_number") %>% 
  left_join(winter_points, by = "Record_number") %>% 
  ggplot(aes(x = Rs_winter, y = Rh_annual)) +
  theme_cowplot() + 
  geom_smooth(method='lm', color = "black", se = FALSE)-> p

p + geom_point(aes(color = map_winter), size = 3) + 
  scale_color_viridis(option = "mako", begin = 1, end = 0.4) -> p1

p + geom_point(aes(color = mat_winter), size = 3)  + 
  scale_color_viridis(option = "inferno", begin = 1, end = 0.4) -> p2

p + geom_point(aes(color = precip_seasonality), size = 3)  + 
  scale_color_viridis(option = "mako", begin = 1, end = 0.4) -> p3

p + geom_point(aes(color = temp_seasonality), size = 3) + 
  scale_color_viridis(option = "inferno", begin = 1, end = 0.4) -> p4

plot_grid(p1, p2, p3, p4)

```

Some studies report mat and map in SRDB, let's check how similar these are to WorldClim data
```{r checks, echo=FALSE}

# Some studies report mat and map in SRDB, let's check how similar these are to WorldClim data
srdb_winter %>% 
  dplyr::select(Record_number, Study_number, MAT, MAP) %>% 
  left_join(winter_points, by = c("Record_number", "Study_number")) -> d

ggplot(d, aes(x = MAT, y = mat_winter)) +
  geom_abline(intercept = 0) +
  geom_point() -> mat

ggplot(d, aes(x = MAP, y = map_winter)) +
  geom_abline(intercept = 0) +
  geom_point() -> map

plot_grid(mat, map)

```

```{r linear-model}

srdb_winter %>% 
  dplyr::select(Record_number, Study_number, Rs_winter, Rh_annual, Biome, Leaf_habit) %>% 
  left_join(seasonality, by = "Record_number") %>% 
  left_join(winter_points, by = "Record_number") -> fulldata

lm(Rh_annual ~ Rs_winter + mat_winter + map_winter + temp_seasonality + precip_seasonality + Leaf_habit,
   data = fulldata) -> lm

summary(lm)
```

