---
title: "Science Question: Can winter Rs a proxy for annual Rh? Can this be scaled globally?"
author: "Stephanie Pennington"
date: "9/29/2021"
output: html_document
---

Annual soil respiration (Rs) is composed of autotropic (Ra) and heterotrophic (Rh) respiration. 

**Hypothesis 1:** In the winter, plants are dormant and produce negligible respiration. Therefore we hypothesize that wintertime Rs is mainly comprised of microbial activity (Rh). 

_How do we test H1?_ 
- COSORE? Need sites with wintertime Rs and Rh measurements
- SRDB? What studies report Rs_winter and separate Rh and Ra annually. Some may have reported wintertime Rh but not entered into SRDB
- Lit searching?
- AmeriFlux? Sites report ER (total ecosystem respiration); if in winter ER gets close to Rs (from SRDB?) that's suggestive, needs more research

**Hypothesis 2:** Wintertime Rs (equals Rh, if H1 is true) can be used as a proxy for annual Rh. This relationship could allow for biome-scale estimates of Rh.

_How do we test H2?_
- SRDB: look at studies that report Rs_winter and Rh_annual
- COSORE?

```{r setup, include = FALSE}
library(dplyr)
library(readr)
library(tidyr)
library(DT)

# We want to load in the SRDB and search for relevant studies
srdb_data <- read_csv("../Data/srdb-data.csv")

vars <- c("Rs_winter", "Rh_annual")
```

```{r , inclue = FALSE}
# How many Rs_winter observations are there in SRDB?
winter_rs <- nrow(drop_na(srdb_data, Rs_winter))
    
# Also how many Rs_winter observations do we have that also have an Rh_annual value? 
srdb_data %>% 
    select(Record_number, Study_number, Study_midyear, Biome, Rs_winter, Rs_annual, Rh_annual) %>% 
    drop_na(all_of(vars)) -> srdb_winter

rsrh_obs <- nrow(srdb_winter)
rsrh_studies <- length(unique(srdb_winter$Study_number))

# Let's make an interactive table of the winter SRDB entries
datatable(srdb_winter)
    
```

### There are `r winter_rs` `Rs_winter` measurements reported in the SRDB.

### There are `r rsrh_obs` observations in `r rsrh_studies` studies that report both `Rs_winter` and `Rh_annual` reported in the SRDB.

#### Key Terms to Search For
- soil respiration; winter; heterotrophic respiration
